---
title: "TeachingWithMarkdwon"
author: "Andrew Laws"
date: "10/12/2021"
output: html_document
---

### This setup section is where libraries and data are loaded in 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#libraries
library(sf)
library(terra)
library(dplyr)
library(spData)
library(tmap)

#data
data(seine)
```

Learning Objectives:

1. When would you use simplification? Think about what it changes in the data and
   how this would impact maps and analysis.
   
2. What is maintained and what is not maintained during an affine transformation?

   
##Simplification##

- Generalizes vector objects (lines and polygons) by reducing vertexes
- Reduces disk space and speeds up map making
- Used for smaller scales only

Example:
```{r Simplification - Visual}
#View original data
orig <- tm_shape(seine) + tm_lines(col = "black") + tm_layout(title = "Original")
orig

#simplify the gometry
seine_simp = st_simplify(seine, dTolerance = 2000)  # 2000 m

#view new data next to original data
simplify <- tm_shape(seine) + tm_lines(col = "black") + tm_layout(title = "Simplified")

tmap_arrange(orig, simplify, ncol = 2)

#see next chunk for size comparison
```
The geometry has changed some but can you tell how much the data has been simplified? Run the next code chunk to see the disk space reduction

```{r Simplification - Disk Space}
#original disk size
orig.disk <- object.size(seine)

print("Original disk size")
print(orig.disk)

#simplified disk size
simp.disk <- object.size(seine_simp)

print("Simplified disk size")
print(simp.disk)

print("That's a reduction of half the size!")
```

We have a geometry, nz_sfc, that needs to be shifted 3000 meters North and 5000 meters West. How would you type out the vector to perform this shift?
```{r}
#use nz_sfc with a vector to shift
nz_sfc = st_geometry(nz)

#uncomment code below and create shift
nz_shift <- nz_sfc + c(1000000, -5000)

#uncomment code below to map shift
tm_shape(nz_sfc) + tm_polygons(col = "gray") +
 tm_shape(nz_shift) + tm_borders(col = "red") 
```

