---
title: "Lab3"
author: "Andrew Laws"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# libraries
library(spdep)
library(tmap)
library(tmaptools)
library(sf)
library(tidyverse)
library(rgeoda)

# data
aoi <- sf::read_sf("../data/orig/aoi.shp") %>% sf::st_make_valid() %>% 
  sf::st_transform(., "ESRI:102010") #set projection

```

## Task 1
```{r task1, include=FALSE}
# us <- sf::read_sf("../data/orig/County_2010Census_DP1.shp") %>% sf::st_make_valid()
# 
# # subset data to 4 corn statses
# subset <- us %>% filter(., stringr::str_starts(GEOID10, "04") | stringr::str_starts(GEOID10, "08") 
#                      | stringr::str_starts(GEOID10, "49")| stringr::str_starts(GEOID10, "35")) 
#   
# #write to shapefile
# st_write(subset, "../data/orig/aoi.shp", driver="ESRI Shapefile")
# 
# #visualize
# tm_shape(subset) + tm_polygons()
```

## Task 2
```{r task2}
#percentage of households that are single parent
aoi <- aoi %>% mutate(., oneparent = ((DP0130007 + DP0130009)/ DP0130001) * 100) #single male and female parents
                                                                           # over total households
```

## Task 3
```{r task3}
#histogram of variable
hist(aoi$oneparent)
```

## Task 4
```{r task4}
#create bbox to zoom out map
bbox <- st_bbox(aoi)

#get range of values
xrange <- bbox$xmax - bbox$xmin
yrange <- bbox$ymax - bbox$ymin 

#alter corner poitns
bbox[1] <- bbox[1] - (0.25 * xrange) # xmin - left
bbox[3] <- bbox[3] + (0.25 * xrange) # xmax - right
bbox[2] <- bbox[2] - (0.25 * yrange) # ymin - bottom
bbox[4] <- bbox[4] + (0.25 * yrange) # ymax - top

# make a choropleth map
tm_shape(aoi, bbox = bbox) + tm_polygons(col = "oneparent", palette = "-RdBu", n=6, style="jenks", title = "Legend", title.size = 0.5) +
  tm_layout(title = "% of Households Classified as Single Parent", title.size = 2)

```
## Task 5
```{r task5}
#row standardized queen contiguity weights matrix
qn.lw <- poly2nb(aoi, queen = TRUE) %>% nb2listw(., style = "W")

# histogram of neighbors
qn.nbr <- attr(qn.lw$weights, "comp")$d
hist(qn.nbr)

#calculate average number of neighbors
mean_qn.nbr <- qn.nbr %>% as_tibble() %>% 
  summarize(., mean.nbr = mean(value))

print(mean_qn.nbr$mean.nbr)

#create moran plot
qn.moran_plot <- moran.plot(aoi$oneparent, qn.lw)
```

## Task 6
```{r task6}
#convert aoi to poinst
aoi.cent <- st_centroid(aoi)

#get minimum distance between points
thresh <- rgeoda::min_distthreshold(aoi.cent)

#row standardized IDW weights matrix
idw.lw <- dnearneigh(aoi.cent, 0, thresh * 1.25, longlat=FALSE) %>% nb2listwdist(., aoi.cent, type = "idw", style = "W")

# histogram of neighbors
idw.nbr <- attr(idw.lw$weights, "comp")$d
hist(idw.nbr)

#calculate average number of neighbors
mean_idw.nbr <- idw.nbr %>% as_tibble() %>% 
  summarize(., mean.nbr = mean(value))

print(mean_idw.nbr$mean.nbr)

#create moran plot
idw.moran_plot <- moran.plot(aoi.cent$oneparent, idw.lw)
```

## Questions
Questions:
1. Describe in your own words how Moran’s I is calculated

Moran's I is calculated by summing the spatially weighted means for all features and dividing by the total spatial weights.

2. Describe in your own words: what is a spatially-lagged variable?

A spatially-lagged variable is the mean weighted value of a features neighbors and when combined with the features values allow identification of clustering and outlier identification.

3. How does your analysis in this lab (as simple as it is) differ by how you have formalized W (e.g., space,
neighbors) in two different methods? How might it affect analysis?

The contiguity based analysis relies on polygons that have lines that touch to be considered neighbors whereas the IDW method uses distance and centroids, which could create additional neighbors for geometries that don't share common borders. As the weights are row standardized, these additional neighbors could weaken the weight of neighbors present in the contiguity-based weights. 

4. What does it mean if an observation falls in the “H-L” quadrant? Why might it be useful to detect
such occurrences?

This means that the features value is higher than average and is surrounded features with values that are lower than average. Simply put, its a high value surrounded by much lower values and could be considered a spatial outlier.
