---
title: "Lab4"
author: "Andrew Laws"
date: "11/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## libraries
library(spdep)
library(tidyverse)
library(tmap)
library(sf)
library(raster)
library(grid)
library(httr)

## data

#crs
crs <- st_crs("ESRI:102008")

#state scale
counties <- sf::read_sf("../data/County_Boundaries-_Census.shp") %>% st_make_valid() %>% 
  st_transform(., crs=crs)

state <- sf::read_sf("../data/Nebraska_State_Boundary.shp") %>% st_make_valid() %>% 
  st_transform(., crs=crs)

stparks <- sf::read_sf("../data/State_Park_Locations.shp") %>% st_make_valid() %>% 
  st_transform(., crs=crs)

streams <- sf::read_sf("../data/Streams_303_d_.shp") %>% st_make_valid() %>% 
  st_transform(., crs=crs)

#county scale
lanc <- sf::read_sf("../data/lancaster_county.shp") %>% st_make_valid() %>% 
  st_transform(., crs=crs)

dem <- raster::raster("../data/lancaster_dem/lc_dem.tif")

muni <- sf::read_sf("../data/Municipal_Boundaries.shp") %>% st_make_valid() %>% 
  st_transform(., crs=crs)

#second map data

lnk <- filter(muni, NAME == "Lincoln")

#get census tract data from ESRI Living Atlas
url <- parse_url("https://services.arcgis.com/P3ePLMYs2RVChkJx/ArcGIS/rest/services")
url$path <- paste(url$path, "ACS_Poverty_by_Age_Boundaries/FeatureServer/2/query", sep = "/")
url$query <- list(where = "State = 'Nebraska'",
                  outFields = "*",                  
                  returnGeometry = "true",                  
                  f = "json")
request <- build_url(url)

tracts <- st_read(request) %>% dplyr::select(., GEOID: B17020_001E | B17020_002E | geometry) %>% 
  mutate(., pov.perc = B17020_002E/B17020_001E * 100) %>%
  st_transform(., crs=crs)

lnk.tracts <- tracts[lnk,] #intersects with Lancaster county to narrow dataset

#lnk park boundares
lnk.parks <- st_read("https://opendata.arcgis.com/datasets/d325c636525642f2b98bbebd37e22201_0.geojson") %>% #data from Lincoln OpenData
  st_transform(., crs=crs)

#lakes data and subset to Lancaster
lakes <- st_read("https://opendata.arcgis.com/datasets/ed78e82b2727468f91cb3bc1f49f0042_2.geojson") %>% #data from NE OpenData
  st_transform(., crs=crs)

lnk.lakes <- lakes[lnk,] # intersects with Lancaster county
```

```{r state map}
#non family household percentages
counties <- mutate(counties, nfh.perc = NonFamHous/TotalHouse * 100)

st.map <- tm_shape(counties) + tm_polygons(col = "nfh.perc", n=5, style="jenks",
                                           title="Percent", palette="PuBu",
                                           border.col = "yellow", legend.show = FALSE) +
  tm_scale_bar(width = 0.1, position = c("right", "top")) +
  tm_shape(lanc) + tm_borders(col = "red")
st.map
```

```{r county map}
#subset parks, streams, and muni to lanc
l.parks <- st_intersection(stparks, lanc)

l.streams <- st_intersection(streams, lanc)

l.muni <- st_intersection(muni, lanc)

#create map
county.map <- tm_shape(lanc) + tm_borders() + 
  tm_shape(dem, raster.downsample=FALSE) + tm_raster(style= "quantile", n=7, palette="Greys") +
  tm_shape(lanc) + tm_polygons(col = "white", alpha = 0.4) +
  tm_shape(l.muni) + tm_polygons() + tm_text("NAME", size = 0.9, col = "blue") +
  tm_shape(l.parks) + tm_dots(col = "red", size = 1.1, shape = 10) +
  tm_shape(l.streams) + tm_lines(col = "Impairment") +
  tm_compass(position = c("right", "bottom")) +
  tm_layout(legend.show = FALSE, title = "Lancaster County Phenomena", title.position = c("left", "top"), 
            title.color = "black", title.bg.color = "white", title.bg.alpha = 0.8)
county.map
```

```{r combine maps}
#combine maps
county.map
print(st.map, vp=viewport(0.605, 0.85, width = 0.2, height = 0.11))
```

```{r custom map}
tm_shape(lnk) + tm_borders(col = "white") +
  tm_shape(dem) + tm_raster(style= "quantile", n=7, palette="Greys", legend.show = FALSE) + 
  tm_shape(lnk.tracts) + tm_polygons(col = "pov.perc", border.col = "gray", title = "% in Poverty", alpha = 0.9) +
  tm_shape(streams) + tm_lines(col = "blue") +
  tm_shape(lnk.lakes) + tm_polygons(col = "blue", legend.show = TRUE) + 
  tm_shape(lnk.parks) + tm_polygons(col = "#098908") +
  tm_compass(position = c("right", "bottom")) +
  tm_scale_bar(width = 0.1, position = c("left", "bottom")) +
  tm_layout(title = "Poverty and Nature Access", 
            title.color = "black", title.bg.color = "white", title.bg.alpha = 0.8,
            legend.bg.color = "white", legend.bg.alpha = 0.8)

```

## Questions
# 1. Describe your choices in making map 1
I chose gray for the DEM as this is a commonly used color for this. I then chose red targets for park locations since those could be hunting locations. I put the inset map in a location where it wasn't blocking the other features of interest.

# 2. Describe your choices in making map 2
I chose to highlight, at a broad level, poverty and acces to nature around link. I brought in ACS 2019 data with poverty measurements from the ESRI Living Atlas, lakes data from NebraskaMaps.gov, and Lincoln parks from the Lincoln OpenData site, all using web sources. I calculated the % poverty from the ACS data and then plotted the census tracts. I was trying to visualize that higher poverty rates often accompany less access to nature. 

# 3. What did you learn?
I still prefer making maps using a GUI interface. A map requiring and inset is better on a landscape mode map with text filling the whitespace and this is really not feasible to do with tmap or R. Accessing open data was pretty easy programmitcally. 